---
title: "Homework Project 2"
author: "Josh Liu"
format: html
editor: visual
---

## Introduction

The Mandarin Chinese language, with roots that date back thousands of years, is complex but beautiful language. As a logographic language, Chinese characters represent one of the oldest continuously used systems of writing in the world, and is currently used as a communicative medium by almost one fifth of the world’s population. While other languages using phonetic alphabets have undergone radical changes through time, Mandarin Chinese exhibits remarkable continuity which holds profound meaning and tradition for those who speak it. As a native English speaker and a Chinese language student here at the University of Michigan, I have experienced, first-hand, the difference in cognitive processing of symbolic characters versus alphabetic letters, and the difficulties of learning a logographic language. Given the language’s wide usage, lack of research as compared to alphabetic languages, and connection to my personal and academic journey, it is useful to gain a better understanding of it.

Using a lexical decision task, where participants had to discriminate between real Chinese characters and pseudocharacters via button presses, The "Chinese Lexicon Project" provides a valuable database of behavioral responses (reaction times and accuracy) for lexical processing of single Chinese characters. Through analysis of this study's lexical processing data, as well as an alternative study's data on the frequency of Chinese characters, I hope to answer several questions including (1) the relationship between character frequency and lexical decision reaction times, (2) how various lexical properties of characters (e.g., number of strokes, phonetic/semantic radical, number of meanings, etc.) influence lexical decision latencies, (3) the correlation between semantic meaning and response time, and (4) whether or not a character's tone has any impact on it's recognizability or lexical decision reaction time.

## Importing Data

```{r}

library(tidyverse)
library(ggplot2)

clp_raw <- read_csv("/Users/joshliu/Desktop/LING 343/hp2-joshliu17/data/Chinese Lexicon Project Sze et al.csv")
frequency_raw <- read_csv("/Users/joshliu/Desktop/LING 343/hp2-joshliu17/data/Chinese language database _ 中文数据库 - All Characters (Frequency).csv")

```

## Data Cleanup

Here, I cleaned up the Chinese Language Database (Frequency) dataset to make it easier to interprt. Specifically, the descriptions of the columns were  in the first row of the data set rather than being the column names, and the number percentages were written with a comma instead of a period. I renamed the columns, removed unnecessary columns, and replaced the commas with periods.

```{r}

#Renaming columns
frequency_renamed <- frequency_raw %>%
  rename(na = "...1") %>%
  rename(Index = "...2") %>%
  rename(Frequency = "...3") %>%
  rename(General_Standard_Num = "...4") %>%
  rename(Radical = "...5") %>%
  rename(Encounters = "...6") %>%
  rename(Fraction_of_Language = "...7") %>%
  rename(HSK_2.0 = "...8") %>%
  rename(HSK_3.0 = "...9") %>%
  rename(Stroke_Count = "...10") %>%
  rename(Character = "...11") %>%
  rename(Pinyin_Tones = "...12") %>%
  rename(Pinyin_No_Tones = "...13") %>%
  rename(Tone = "...14") %>%
  rename(Meaning = "...15") 

#Removing unnecessary columns
frequency_renamed <- frequency_renamed %>%
  select(-c("na", "Index","HSK_2.0", "HSK_3.0", "Pinyin_No_Tones", "...16"))

#Removing unnecessary rows
frequency_renamed <- frequency_renamed[-c(1, 2), ]

#Replacing commas with periods in the Frequency column
frequency_renamed$Fraction_of_Language <- gsub(",", ".", frequency_renamed$Fraction_of_Language)

```

## Data Dictionaries

Below are the Data Dictionaries for both data sets I will be working with.

| Variable      | Description                                                                                                              |
|---------------------|---------------------------------------------------|
| **Character** | the Chinese character in simplified script                                                                               |
| **Acc**       | mean accuracy                                                                                                            |
| **Ntrials**   | number of participants whose trials were sufficiently reliable to provide the latencies for that item (maximum being 35) |
| **RT**        | mean response time for the item, computed across participants                                                            |
| **SE**        | standard error of the response time for each character                                                                   |
| **SD**        | Standard deviation of the response time for each character                                                               |

: Chinese Lexicon Project

| Variable                 | Description                                                                                                                                                                                                 |
|--------------------------------------|--------------------------------------------------------------|
| **Frequency**            | frequency ranking (from highest to lowest)                                                                                                                                                                  |
| **General_Standard_Num** | character's index number according to the Table of General Standard Chinese Characters, the current standard list of 8,105 Chinese characters published by the government of the People's Republic of China |
| **Radical**              | character's radical, a graphical component of a Chinese character that indicates the character's meaning or pronunciation                                                                                   |
| **Encounters**           | number of times the character is encountered in the corpus                                                                                                                                                  |
| **Fraction_of_Language** | how much of the language is made up by the character; the number of times the character is encountered in the corpus in relation to total amount of Chinese characters                                      |
| **Stroke_Count**         | number of strokes in the character                                                                                                                                                                          |
| **Character**            | Chinese character                                                                                                                                                                                           |
| **Pinyin_Tones**         | character's pinyin (romanized spelling for chinese characters) with tone marks                                                                                                                              |
| **Tone**                 | character's tone (first through fourth)                                                                                                                                                                     |
| **Meaning**              | character's semantic meaning(s)                                                                                                                                                                             |

: Chinese Frequency Database

## Analysis

To investigate the relationship between character frequency and lexical decision reaction times, I have to combine the two datasets (1) containing the lexical decision times, and (2) containing character frequency, using a join. 

```{r}

#Joining the two data sets
combined_data <- clp_raw %>%
  left_join(frequency_renamed)

```

Next, I decided to creat a regression model to view the correlation between the two variables. 

```{r}

#creating a regression model
ggplot(combined_data, aes(x = RT, y = Frequency)) +
  geom_point(col = "violet") + 
  labs(title = "Mean Response Time vs. Frequency",
       x = "Mean Response Time",
       y = "Frequency") +
  xlim(c(400, 900)) +
  geom_smooth()

```
The regression model above clearly shows a positive relationship between a character's Mean Response Time and Frequency in the corpus—as a character's frequency ranking rises, it's mean response time among participants also increases. This suggests the less frequent a character is, the less identifiable it is. This notion of recognizability is not a novel discovery, but suggests that the study is accurate in this sense.


















